Palomar 5 Gaps

### 2024-10-28 paper/main.tex
more discussion points
discussion and trying to handle Pal 5's mass

    At their computed positions 5 Gyr ago, each globular cluster was assigned 100,000 particles sampled from a Plummer distribution to represent its stellar content. For each set of initial conditions, the particle positions were resampled to account for uncertainties in mass, which is an improvement from \citet{2023A&A...673A..44F}. Since globular clusters are older than 8 Gyr, their most massive stars have already gone supernova, leaving primarily low-mass stars, typically under one solar mass (\@\salvatore{Citation? Paolo knows this stuff well.}). Given that clusters usually have masses between $10^5-10^6$ solar masses, 100~000 particles is sufficient to represent the distribution of tidal debris.

@@

    % We initially sought an automated, quantitative approach from signal and image processing that could reliably detect all gaps and evaluate their significance. Unfortunately, despite extensive efforts, an ideal method remained elusive. As a result, we opted for a simpler approach: a combination of visual inspection aided by comparative analysis.

@@

    \begin{itemize}
      \item Would these gaps be observable? We are in the max info limit, what about observational difficulties?
      \item the detected track of Palomar 5 is much shorter than my simulations. Does this mean cluster gaps sould excedingly rare?
      \item What about Carlberg's Pal 5 gap paper? 
      \item What about Bonaca's Pal 5 gap paper?
      \item future works?
      \item bars substructure?
      \item What else?
      \begin{itemize}
        \item Doing a complete census of all globular clusters?
        \item doing a compelte census of all observed streams? 
        \item comparing to dark matter subhaloes?
        \end{itemize}
      \end{itemize}

>>>>

    At their computed positions 5 Gyr ago, Palomar 5 was assigned 100,000 particles sampled from a Plummer distribution to represent its stellar content. For each set of initial conditions, the particle were resampled based on uncertainties in mass, which is an improvement from \citet{2023A&A...673A..44F}.

    We use a mass estimate for Palomar 5 taken from \citet{2021MNRAS.505.5957B}, $1.34\pm0.24\times10^{4}M_{\odot}$--which is a contested value.

    \salvatore{I'm a bit lost on trying to reconcile the differences between Baumgardt's estiamte and \citet{2017ApJ...842..120I}'s estimate in sections 5.2 and 5.3. It looks like 5.2 is based on counts and then supplementing the completness function with dNdM \citet{2001AJ....122.3231G} and demonstrating that Salpeter would overestimate the mass. I think 5.3 is showing that an MCMC was used to add more context to the estimate from 5.3.}

    \salvatore{\textbf{If Salpeter over estimates the number of low mass stars}, then wouldn't this appear in the velocity dispersion profile? or wait, is \citet{2017ApJ...842..120I} fitting the density for the King, supplementing missing stars by a Mass function from \citet{2001AJ....122.3231G}, and finding the total mass this way? i.e. no velocity dispersion was investigated? I feel like viewing the velocity dispersion directly would provide a good estiamte for the dynamical mass, no?}

    \salvatore{I am lost in Baumgardt's papers trying to flush out how a MF was used to provide their mass estimate. It looks like this information is within Sollima 2017. Need to revisit this. }

    \salvatore{What I say will depend. I think using \citet{2021MNRAS.505.5957B}'s mass estimate is still ok because it is similar in order of magnitude to cluster + stream from \citet{2017ApJ...842..120I}. }

@@

@@

  In this work, we quantified the number of gaps expected in Palomar 5's stream from the 165 Galactic globular clusters using a Monte Carlo-style experiment. We performed 50 iterations, varying the orbits based on the uncertainties in their observed proper motions, radial velocities, and distances. Our findings are as follows:

  \begin{itemize}
    \item \textbf{Orbital Overlap}: clusters with orbital energies and angular momenta that overlap with Palomar 5's orbit can impact its stream.
    \item \textbf{Mass Threshold}: Perturbing clusters have masses above $10^5 M_\odot$; no perturber has a mass smaller than this.
    \item \textbf{Galactic radius}: Impacts preferentially occur at shorter galactic radii.
    \item \textbf{Momentum and Frequency Change}: Gap creation is linked to the momentum imparted to the stream during the flyby, yet not entirely. As showcased by \citet{2016MNRAS.457.3817S}, the change in orbital frequencies induced by the flyby must be greater than the spread in frequencies present in the stream. Some impacts imparted significant momentum changes that did not lead to gap growth for this reason.
  \end{itemize}

  Finally, we present a gap creation rate as conceptualized by \citet{2012ApJ...748...20C}, which is the number of gaps normalized to the age and length of the stream. This rate is: 0.009~[km~s$^{-1}$~kpc$^{-2}$], or 0.01~[Gyr$^{-1}$~kpc$^{-1}$].

  Now, \textit{is this consistent with reality?} Can globular clusters explain the observed gaps in Palomar 5? Currently, the results on the number of gaps in the stream are debated. \citet{2012ApJ...760...75C} and \citet{2020ApJ...889...70B} attempted to quantify the stream's density along its width, isolating the stream from background and foreground contamination. \citet{2020ApJ...889...70B} reported the existence of two potential gaps, whereas \citet{2012ApJ...760...75C} identified five gaps with a 99~\% detection confidence, leading to a gap creation rate of 0.17~Gyr$^{-1}$~kpc$^{-1}$. If true, this rate is too high to be explained by Galactic globular clusters.

  On the other hand, for a stream length of 10~kpc reported from \citet{2020ApJ...889...70B} and given our gap creation rate, in 5~Gyr we would expect to see about 0.5 gaps, in 10~Gyr's we would expect 1. Therefore, finding two gaps from globular clusters would not be impossible.

    \begin{itemize}
      \item Would these gaps be observable? We are in the max info limit, what about observational difficulties?
      \item the detected track of Palomar 5 is much shorter than my simulations. Does this mean cluster gaps sould excedingly rare?
      \item future works?
      \item bars substructure?
      \item link to tstrippy? open source
      \item add the new simulations to etidal public database?
      \item the new proposed instrument and survey by Conroy and Bonaca to get improve stream data? https://via-project.org/. How to cite or include in the discussion?
      \item \textbf{What else?}
      \item Future potential works
      \begin{itemize}
        \item mock observations to build confusion matrix for gaps true/false positive/neagtive detections. 
        \item Doing a complete census of all globular clusters?
        \item doing a compelte census of all observed streams? 
        \item comparing to dark matter subhaloes?
        \end{itemize}
      \end{itemize}
